---
title: "Partisan Dehumanization Survey Experiment"
output: html_notebook
---

In this project I clean and transform data from Qualtrics. I ran a 3x2 design in 
which I had treatment/control condition (2) and vignette/graph/image stimulus (3).
The goal of the experiment to see which manipulation worked best (vignette/graph/image).
The treatment and control also differed by the partisanship of the respondent. For example, A
Republican assigned to the vignette condition, would recieve a different vignette 
than a Democrat assigned to the same condition. Independents were randomly assigned 
to be treated as either Democrats or Republicans. This effectively turns the experiment 
into a 2x2x3: Republican/Democrat, treatment/control and vignette/graph/image.

As a result, the raw data is 'scattered' as Qualtrics simply creates a unique column 
for each condition. In this script I clean, condense, and transform the data so 
that it is suitable for statistical analysis. 

```{r}
rm(list = ls())

setwd("C:/Users/cayde/OneDrive/Graduate Studies/Research/Affective Polarization/Data")
pilot_data_raw <- read.csv("Dehumanization Pilot Data (Master).csv", header=FALSE, comment.char="#")

library(dplyr)
```

Here I do some very basic removal of columns that do not contain relevant data.

```{r}

#Removing unnecessary rows and columns
pilot_data_draft <- pilot_data_raw #Creating a duplicate of the data to preserve the original 
pilot_data_draft <- pilot_data_raw[- c(1,2,3),] #removing question wording rows
pilot_data_draft <- pilot_data_draft[pilot_data_draft$V18 != "I do not wish to participate in the survey",] #removing people who opted out of the survey
pilot_data_draft <- pilot_data_draft[,-c(1:18)] #removing irrelevant meta data
```

All the variables in qualtrics are named "V1", "V2", etc. I rename some of the 
basic demographic info
```{r}


#Renaming variables
pilot_data_draft <- rename(pilot_data_draft, "sex" = "V19")
pilot_data_draft <- rename(pilot_data_draft, "age" = "V20")
pilot_data_draft <- rename(pilot_data_draft, "race" = "V21")
pilot_data_draft <- rename(pilot_data_draft, "education" = "V22")
pilot_data_draft <- rename(pilot_data_draft, "PID" = "V23")
```
Here I simply create new variables based on teh values of other variables
```{r}
#Converting Party Identification to numeric
pilot_data_draft$PID7 <- recode(pilot_data_draft$PID, "Strong Democrat" = 1,
                                "Democrat" = 2,
                                "Lean Democrat" = 3,
                                "Independent" = 4,
                                "Lean Republican" = 5,
                                "Republican" = 6,
                                "Strong Republican" = 7)

table(pilot_data_draft$PID, pilot_data_draft$PID7) #checking the re-coding

#Creating new binary variable that measures partisanship
pilot_data_draft$partisan <- 0
pilot_data_draft$partisan[pilot_data_draft$PID != "Independent"] <- 1
```
This is where things get a little bit tricky. I have 4 variables each of which 
are measured on a scale of 1-100. However, because there are two versions of the 
survey (Republican/Democrat) each of these four questions gets duplicated. For
example, V24 and V34 are the same question, but V24 was the one the Republicans
answered and V34 was the one Democrats answered. I thus need to merge these 
columns into a new variable. 
```{r}
#Getting Dehumanization, Feeling, and Meta ratings for Democrats and Republicans 
pilot_data_draft$ascent_D <- as.numeric(paste(pilot_data_draft$V24, pilot_data_draft$V34))
pilot_data_draft$ascent_R <- as.numeric(paste(pilot_data_draft$V25, pilot_data_draft$V33))

pilot_data_draft$therm_D <- as.numeric(paste(pilot_data_draft$V27, pilot_data_draft$V35))
pilot_data_draft$therm_R <- as.numeric(paste(pilot_data_draft$V26, pilot_data_draft$V36))
```
Same process as before but the variable is a likert scale rather than a 1-100
so I need to recode it before converting it to numeric.
```{r}
pilot_data_draft$meta_prejudice <- paste(pilot_data_draft$V28, pilot_data_draft$V37)
pilot_data_draft$meta_dehum <- paste(pilot_data_draft$V29, pilot_data_draft$V38)
```

Just converting character variables to numeric so I can combine them into an index
```{r}
#Re-coding individual batter items: Political Threats, harassment, and violence
pilot_data_draft$political_threats <- recode(pilot_data_draft$V30, "1 (never)" = 1,
                                             "2" = 2,
                                             "3" = 3,
                                             "4 (always)" =4)

pilot_data_draft$political_harassment <- recode(pilot_data_draft$V31, "1 (never)" = 1,
                                                "2" = 2,
                                                "3" = 3,
                                                "4 (always)" =4)

pilot_data_draft$political_violence <- recode(pilot_data_draft$V32, "1 (not at all)" = 1,
                                              "2" = 2,
                                              "3" = 3,
                                              "4" = 4,
                                              "5 (a great deal" = 5)

#Combining items into an index
pilot_data_draft$toxic <- pilot_data_draft$political_violence + pilot_data_draft$political_harassment + pilot_data_draft$political_threats
```
Here I address the problem of Qualtrics seperating all the data. I take advantage 
of the fact the data is in character format, and the empty cells simply have "" as 
a value. Using the paste command I can merge all the data. I end up with a column
that as three letters, one letter for each part of 2x2x3 design. 
```{r}
#Creating experiment assignment based on what version survey the respondent took
pilot_data_draft$experiment <- trimws(paste(pilot_data_draft$V57, pilot_data_draft$V58,  #Independents assigned the D/R) vignette
                                            pilot_data_draft$V64, pilot_data_draft$V68,  #Democrats and Republicans assigned to their Vignette
                                            pilot_data_draft$V63, pilot_data_draft$V67,  #Democrats and Republicans assigned to their Image
                                            pilot_data_draft$V54, pilot_data_draft$V55,  #Independents assigned the D/R Image 
                                            pilot_data_draft$V65, pilot_data_draft$V69,  #Democrats and Republicans assigned to their Image
                                            pilot_data_draft$V60, pilot_data_draft$V61)) #Independents assigned to the Republican or Democratic Graph

pilot_data_draft$experiment <- recode(pilot_data_draft$experiment, "FL_177" = "RCG",     #Due to Qualtrics, the control graph condition needs to be renamed
                                      "FL_181" = "DCG",
                                      "FL_186" = "RCG",
                                      "FL_190" = "DCG")
```

I create three new columns based off of the three letter assignment column. This
way I can more easily filter and subset the data when I analyze it. 

```{r}
#Separating the experiment condition into three new columns. Each letter in the experiment assignment corresponds to a certain condition 
pilot_data_draft$version[sapply(pilot_data_draft$experiment, function(x){"I" %in% unlist(strsplit(x, split = ""))})] <- "Image"
pilot_data_draft$version[sapply(pilot_data_draft$experiment, function(x){"V" %in% unlist(strsplit(x, split = ""))})] <- "Vignette"
pilot_data_draft$version[sapply(pilot_data_draft$experiment, function(x){"G" %in% unlist(strsplit(x, split = ""))})] <- "Graph"

pilot_data_draft$target_party[sapply(pilot_data_draft$experiment, function(x){"R" %in% unlist(strsplit(x, split = ""))})] <- "Republican"
pilot_data_draft$target_party[sapply(pilot_data_draft$experiment, function(x){"D" %in% unlist(strsplit(x, split = ""))})] <- "Democrat"

pilot_data_draft$condition[sapply(pilot_data_draft$experiment, function(x){"T" %in% unlist(strsplit(x, split = ""))})] <- "Treatment"
pilot_data_draft$condition[sapply(pilot_data_draft$experiment, function(x){"C" %in% unlist(strsplit(x, split = ""))})] <- "Control"
```

Here is some code where I rename the remaining variables. I don't have an immediate 
use for these for my analysis so I leave them commented out for now. 
```{r}
#Renaming remaining Questions
# pilot_data_draft <- rename(pilot_data_draft, "trump_sat" = "V39")
# pilot_data_draft <- rename(pilot_data_draft, "trump_qual" = "V40")
# pilot_data_draft <- rename(pilot_data_draft, "trump_pres" = "V41")
# pilot_data_draft <- rename(pilot_data_draft, "hill_sat" = "V42")
# pilot_data_draft <- rename(pilot_data_draft, "hill_qual" = "V43")
# pilot_data_draft <- rename(pilot_data_draft, "hill_pres" = "V44")
# 
# pilot_data_draft <- rename(pilot_data_draft, "refuse_mask" = "V45")
# pilot_data_draft <- rename(pilot_data_draft, "misinfo" = "V46")
# pilot_data_draft <- rename(pilot_data_draft, "downplay_die" = "V47")
# pilot_data_draft <- rename(pilot_data_draft, "le_just" = "V48")
# pilot_data_draft <- rename(pilot_data_draft, "protestor_hurt" = "V49")
# pilot_data_draft <- rename(pilot_data_draft, "protestor_pepper" = "V50")
```
Finally I remove all the old or transformed columns. Since I gave meaningful names
to all my new variables I simply check to see which variables have a capital "V" 
in them. All the "V1" "V2" ... variables are thus removed. I then save my data to
a file on my computer. 
```{r}
#Removing old variables that have been transformed or that are not needed for final dataset 
pilot_data_draft <- pilot_data_draft[,sapply(names(pilot_data_draft), function(x){!"V" %in% unlist(strsplit(x, split = ""))})]

write.csv(pilot_data_draft, file = "C:/Users/cayde/OneDrive/Data Science/R Projects/Data Transformation/Pilot Data Cleaned.csv")
```