---
title: "Gender & Political Ambition (Data Cleaning)"
output: html_notebook
---

```{r setup}
rm(list = ls())

library(dplyr)
library(Metrics)
library(ggplot2)

setwd("C:/Users/cayde/OneDrive/Data Science/R Projects/Data Transformation/Cleaning Political Ambition")
pa_data_og <- read.csv("Political Ambition (Dirty).csv")
```

Some basic data cleaning. I remove rows and columns with metadata that I don't 
need. I also identify survey IDs that have a lot of missing data.
```{r}

#Cleaning Data
pa_data <- pa_data_og
pa_data <- pa_data[-c(1:2), -c(1:8, 10:17)]

no_response <- c("659448", "762817", "391501", "802419", "759285", "345561", 
                 "558169", "523583", "706479", "462932", "100319", "338725")
pa_data <- pa_data[!pa_data$SurveyID %in% no_response,]

```

My data is currently in the form of characters. Here I recode the data to either
factor or numeric in order to capture the logical order of the data and make it
suitable for analysis. 

```{r}

#Transforming Data

names(pa_data)

pa_data$Gender <- recode_factor(pa_data$Gender,
                             "Female" = 0,
                             "Male" = 1)

pa_data$College <- recode_factor(pa_data$Education,
                              "Some high school" = 0,
                              "High school graduate" = 0,
                              "Some college" = 0,
                              "2 year degree" = 0,
                              "4 year degree" = 1,
                              "Master's degree" = 1,
                              "Doctorate degree" = 1)

pa_data$Children <- recode_factor(pa_data$Children,
                               "No" = 0,
                               "Yes" = 1)    

pa_data$Income <- recode(pa_data$Income,
                             "Less than $10,000" = 1,
                             "$10,000 - $19,999" = 2,
                             "$20,000 - $29,999" = 3,
                             "$30,000 - $39,999" = 4,
                             "$40,000 - $49,999" = 5,
                             "$50,000 - $59,999" = 6,
                             "$60,000 - $69,999" = 7,
                             "$70,000 - $79,999" = 8,
                             "$80,000 - $89,999" = 9,
                             "$90,000 - $99,999" = 10,
                             "$100,000 - $149,999" = 11,
                             "More than $150,000" = 12)

for(i in which(names(pa_data) == "H1.Flattery"):which(names(pa_data) == "Competition")){
  pa_data[i] <- recode(pa_data[,i],
                           "Strongly disagree" = 1,
                           "Disagree" = 2,
                           "Neither agree nor disagree" = 3,
                           "Agree"  = 4,
                           "Strongly agree" = 5)
}

table(pa_data$Interest.Politics)

pa_data$Interest.Politics <- recode(pa_data$Interest.Politics,
                                        "Not interested at all" = 1,
                                        "Slightly interested" = 2,
                                        "Interested" = 3,
                                        "Very interested"  = 4)

pa_data$Discuss.Politics <- recode(pa_data$Discuss.Politics,
                                       "Never" = 1,
                                       "Occasionally" = 2,
                                       "Sometimes" = 3,
                                       "Frequently"  = 4)            

pa_data$Ideology <- recode(pa_data$Ideology,
                               "Very Liberal" = 1,
                               "Liberal" = 2,
                               "Somewhat liberal" = 3,
                               "Independent"  = 4,
                               "Somewhat conservative" = 5,
                               "Conservative" = 6,
                               "Very conservative" = 7)  

pa_data$PID <- recode(pa_data$PID,
                          "Democrat" = 1,
                          "Independent" = 2,
                          "Other" = 2,
                          "Republican" = 3)

pa_data$Political.Ambition <- recode(pa_data$Political.Ambition,
                                         "Never thought about running" = 1,
                                         "Considered running" = 2,
                                         "Seriously considered running" = 3,
                                         "Ran for office" = 4)

pa_data$Qualification <- recode(pa_data$Qualification,
                                    "Not at all qualified" = 1,
                                    "Somewhat qualified" = 2,
                                    "Qualified" = 3,
                                    "Very qualified" = 4)

pa_data$Winning <- recode(pa_data$Winning,
                              "Very unlikely" = 1,
                              "Unlikely" = 2,
                              "Likely" = 3,
                              "Very likely"  =4)

pa_data$Encourage.Run <- recode(pa_data$Encourage.Run,
                                    "Never" = 1,
                                    "Occasionally" = 2,
                                    "Sometimes" = 3,
                                    "Frequently" = 4)

pa_data$Job.Enjoyment <- recode(pa_data$Job.Enjoyment,
                                    "Dislike a great deal" = 1,
                                    "Dislike somewhat" = 2,
                                    "Neither like nor dislike" =3,
                                    "Like somewhat" = 4,
                                    "Like a great deal" = 5)

pa_data$Marital.Status <- recode(pa_data$Marital.Status,
                                     "Married" = 1,
                                     .default = 0)

pa_data$Ethnicity <- recode(pa_data$Ethnicity,
                                "White" = 1,
                                .default = 0)
```

Some respondents mistakenly put their age when they were asked
for their birth year. I thus have some respondents with birthyear and some with 
age. In order to make the data consistent, I convert the birth year from age
by subtracting the it from the year in which the survey was taken. I only do this
for values over 1000 however, as not to mess up the respondents who had listed 
their age. 

```{r}
#Cleaning Age
##############

#For those that put their birth year (numbers over 1000) I am subtracting the 
#year they took the survey to get their age 
pa_data$Age <- as.numeric(pa_data$Age)
pa_data$Age[pa_data$Age > 1000 & !is.na(pa_data$Age)] <- 2021 - pa_data$Age[pa_data$Age > 1000]
```

Much of the data I have is responses to questions about personality. These 
personality questions need to be combined to form an index or a battery. Using 
the column names and pattern recognition I subset the data by personality trait.
Each personality trait (Honesty, Agreeableness, etc.) has its own dataframe. 
Some of the personality questions are also 'reverse-coded' so that I need to 
flip the question before combining it into a battery.

```{r}
#Battery Creation
#################

#Subsetting the data to transform variables 
honesty <- pa_data[grepl("H\\d", names(pa_data))]
agreeableness <- pa_data[grepl("A\\d", names(pa_data))]
narcissism <- pa_data[grepl("N\\d", names(pa_data))]
machiavellianism <- pa_data[grepl("M\\d", names(pa_data))]
political_ambition <- pa_data[,c('Political.Ambition', "Qualification", "Winning", "Job.Enjoyment")]


#Changing the values for questions that are reverse coded
reverse_col_honesty <- c(2,4,5,7,8,10)
reverse_col_agreeableness <- c(2,3,4,10)

honesty[,reverse_col_honesty] <- 6 - honesty[,reverse_col_honesty]
agreeableness[,reverse_col_agreeableness] <- 6 - agreeableness[,reverse_col_agreeableness]
```

After subsetting the data and reverse coding the items, I have missing data. 
Because I have already ommitted the egregious cases, many of these respondents
are just missing a question or two. Because the batteries are a combination of 
questions, I will not be able to get measurements on a personlity trait if one
item from that battery is missing. In order to preserve data I impute the values.

To do this, I look at each personality trait separately taking advantage of the 
subsets I created earlier. For each respondent with missing data, I take the 
average of their responses, and use that average to fill in the missing data. 

```{r}

############################
#imputing missing variables
############################

sapply(pa_data, function(x){sum(is.na(x))}) #checking to see the missing data

#Personality Traits
###################

#Finds all the rows with NA values and calculates the mean of the row and rounds it
missing_data_honesty <- round(rowMeans(honesty[is.na(rowSums(honesty)),], na.rm = TRUE))

#Using the previous vector, goes through all the rows with NA values and replaces
#all NAs with the mean of the row 
for(row in names(missing_data_honesty)){
  honesty[which(row.names(honesty) == row),][is.na(honesty[which(row.names(honesty) == row),])] <- missing_data_honesty[row]
}

#Repeat this for the other variables

#Agreeableness
missing_data_agreeableness <- round(rowMeans(agreeableness[is.na(rowSums(agreeableness)),], na.rm = TRUE))

for(row in names(missing_data_agreeableness)){
  agreeableness[which(row.names(agreeableness) == row),][is.na(agreeableness[which(row.names(agreeableness) == row),])] <- missing_data_agreeableness[row]
}

#Narcissism
missing_data_narcissism <- round(rowMeans(narcissism[is.na(rowSums(narcissism)),], na.rm = TRUE))

for(row in names(missing_data_narcissism)){
  narcissism[which(row.names(narcissism) == row),][is.na(narcissism[which(row.names(narcissism) == row),])] <- missing_data_narcissism[row]
}

#Machiavellianism
missing_data_machiavellianism <- round(rowMeans(machiavellianism[is.na(rowSums(machiavellianism)),], na.rm = TRUE))

for(row in names(missing_data_machiavellianism)){
  machiavellianism[which(row.names(machiavellianism) == row),][is.na(machiavellianism[which(row.names(machiavellianism) == row),])] <- missing_data_machiavellianism[row]
}

```

I have 13 missing respondents for age. Age is just one of the many variables in 
my model and is mainly used as a control variable or covariate. Because my data
size is small I decide to impute missing age with the average age. 

```{r}
#Imputing Age

pa_data$Age[is.na(pa_data$Age)] <- round(mean(pa_data$Age, na.rm = TRUE))
```

Now that I have imputed data for missing values, I can create new variables in 
the main data set for the personality variables. 
```{r}

#Battery Creation

pa_data$honesty <- rowSums(honesty)
pa_data$agreeableness <- rowSums(agreeableness)
pa_data$machiavellianism <- rowSums(machiavellianism)
pa_data$narcissism <- rowSums(narcissism)
pa_data$political_ambition <- rowSums(political_ambition)

sapply(pa_data, function(x){sum(is.na(x))})

```

Now that the character data has been converted into factors or numbers, and the 
missing data has either been dropped or imputed, the data set is now ready for
statistical analysis. 
